\myTitle

### 2024-11-06 tex/supplemental-figures.tex
Eric's near-final edits

%%%%%%%%%%%%%%%%%%%%%% 
\begin{figure}
\begin{center}
\includegraphics[width=0.8\textwidth]{images/ass-table-80-crop.pdf}
\end{center}
\caption[Assignment table for fish with scaled likelihood $ > 0.8$]{\footnotesize Assignment table
like that in Fig.~\ref{fig:gsi}b in the paper, but constrained so that only fish assigning
to a reporting unit with scaled likelihood greater than 0.8 are included.}
\label{fig:eighty}
\end{figure}
%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%% 
\begin{figure}
\begin{center}
\includegraphics[width=0.8\textwidth]{images/rosa-gsi-table-crop.pdf}
\end{center}
\caption[Assignment table by RoSA genotype]{\footnotesize Assignment table
like that in Fig.~\ref{fig:gsi}b in the paper, but with numbers according to genotypes
at the RoSA. In each cell, the top left entry gives the number of EE (early run allele homozygotes)
genotypes, the middle entry is the number of EL genotypes (heterozygotes), and the bottom
right entry gives the number of LL (late-run homozygote) genotypes. }
\label{fig:rosa-gsi}
\end{figure}
%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%
\begin{figure}
\includegraphics[width=\textwidth]{images/wrap-candi.pdf}
\caption[
	61 candidates for primer design for winter-run-associated polymorphisms
]{
	\footnotesize Genomic positions and values of $|d|$ (absolute difference in
	allele frequency between winter run and non-winter run fish of the CCV)
	for the 61 candidate SNPs (in red) to design for winter-run-associated polymorphisms
	(WRAPs).  Other sites are shown in black.  The $x$-axis shows position on
	each chromsome in the Otsh\_v1.0 assembly.  The chromsome name is indicated in the facet headers by
	RefSeq name
}
\label{fig:wrap-candi}
\end{figure}

>>>>

%%%%%%%%%%%%%%%%%%%%%% 
\begin{figure}
\begin{center}
\includegraphics[width=0.8\textwidth]{images/ass-table-80-crop.pdf}
\end{center}
\caption[Assignment table for fish with scaled likelihood $ > 0.8$]{\footnotesize Assignment table
like that in Figure~\ref{fig:gsi}b in the paper, but constrained so that only fish assigning
to a reporting unit with scaled likelihood greater than 0.8 are included.}
\label{fig:eighty}
\end{figure}
%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%% 
\begin{figure}
\begin{center}
\includegraphics[width=0.8\textwidth]{images/rosa-gsi-table-crop.pdf}
\end{center}
\caption[Assignment table by RoSA genotype]{\footnotesize Assignment table
like that in Figure~\ref{fig:gsi}b in the paper, but with numbers according to genotypes
at the RoSA. In each cell, the top left entry gives the number of EE (early run allele homozygotes)
genotypes, the middle entry is the number of EL genotypes (heterozygotes), and the bottom
right entry gives the number of LL (late-run homozygote) genotypes. }
\label{fig:rosa-gsi}
\end{figure}
%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%
\begin{figure}
\includegraphics[width=\textwidth]{images/wrap-candi.pdf}
\caption[
	61 candidates for primer design for winter-run-associated polymorphisms
]{
	\footnotesize Genomic positions and values of $|d|$ (absolute difference in
	allele frequency between winter run and non-winter run fish of the CCV)
	for the 61 candidate SNPs (in red) to design for winter-run-associated polymorphisms
	(WRAPs).  Other sites are shown in blue.  The $x$-axis shows position on
	each chromsome in the Otsh\_v1.0 assembly.  The chromsome name is indicated in the facet headers by
	RefSeq name
}
\label{fig:wrap-candi}
\end{figure}

### 2024-11-06 tex/supplemental-text.tex
Eric's near-final edits

More specifically, we calculated allele frequencies for 16 winter-run fish and 84 non-winter-run
fish from the  \citet{thompson2020complex} variant data VCF files using ANGSD version~0.921.
Sites were retained if at least 62.5\% of samples in each group had read data (10 of 16 for the
winter run and 50 of 80 for the non-winter run), resulting in 7,295,001 SNPs for downstream
analysis.  We first investigated the distribution throughout the genome of $|d|$, the absolute difference
of the alternate allele frequency between the two groups  (Fig.~\ref{fig:wrap-absdiff}). This revealed
that many loci had large values of $|d|$, but there were several
regions in the genome, in particular, with prominent peaks in allele frequency difference and one
or more SNPs apparently fixed for alternate alleles between the two sample groups.

To leverage information from multiple SNPs to identify regions in the genome with large
allele frequency differences, we calculated the fraction of SNPs with $|d| > 0.5$ that also
had $|d| > 0.9$ within non-overlapping 100 kb sliding windows throughout the genome. This metric indicated
several prominent peaks.  We focused on all of those
100 kb windows in which more than 12.4\% of sites with $|d|>0.5$ also had $|d|>0.9$ which were
also adjacent to at least one window in which $>10\%$ of sites with $|d|>0.5$ also had $|d|>0.9$ 
(Fig.~\ref{fig:wrap-slide-window}).  Within the windows found on four different chromosomes using
the above criteria, we then attempted to design amplicons to type the SNPs at a subset of the sites
within each window.  We chose all sites with $|d|>0.975$, as well as the 8 SNPs on each chromosome
with the highest values of $|d|$, yielding 61 candidate SNPs (Fig.~\ref{fig:wrap-candi}).

@@

%% Note to self.  The next section is documented in the project at:
%% /Users/eriq/Documents/work/assist/anthony_clemento/WRAP-crunching
These 48 amplicons were amplified and sequenced in 192 fish---96 Feather River spring run
and 96 winter run---on a MiSeq sequencer, and the variants were called using
GATK.  The resulting VCF file was used to estimate allele frequency differences between
winter run and feather river spring.  We also processed the sequence data using
microhaplot (\url{https://github.com/ngthomas/microhaplot}) and visually inspected
loci for consisten allele depth ratio and numbers of haplotypes.  We then chose 24 amplicons for further
testing on the basis of allele frequency differences between Feather River spring run and
winter run, number of haplotypes, and ease of scoring.  These 24 markers were typed on
a variety of fish over the course of a year, and we finally chose three to include in
our California Chinook reference baseline: one amplicon on each of chromosomes
8, 12, and 16.  The estimated frequencies of all the alleles present in the reference baseline
in those three amplicons shows that there are not fixed differences between
winter run and all other reporting units (Table~\ref{tab:wrap-freqs}).

>>>>

More specifically, we calculated allele frequencies for 16 winter-run fish and 84 non-winter-run
fish from the  \citet{thompson2020complex} variant data VCF files using ANGSD version~0.921.
Sites were retained if at least 62.5\% of samples in each group had read data (10 of 16 for the
winter run and 50 of 80 for the non-winter run), resulting in 7,295,001 SNPs for downstream
analysis.  We first investigated the distribution throughout the genome of $|d|$, the absolute difference
of the alternate allele frequency between the two groups  (Figure~\ref{fig:wrap-absdiff}). This revealed
that many loci had large values of $|d|$, but there were several
regions in the genome, in particular, with prominent peaks in allele frequency difference and one
or more SNPs apparently fixed for alternate alleles between the two sample groups.

To leverage information from multiple SNPs to identify regions in the genome with large
allele frequency differences, we calculated the fraction of SNPs with $|d| > 0.5$ that also
had $|d| > 0.9$ within non-overlapping 100 kb sliding windows throughout the genome. This metric indicated
several prominent peaks.  We focused on all of those
100 kb windows in which more than 12.4\% of sites with $|d|>0.5$ also had $|d|>0.9$ which were
also adjacent to at least one window in which $>10\%$ of sites with $|d|>0.5$ also had $|d|>0.9$ 
(Figure~\ref{fig:wrap-slide-window}).  Within the windows found on four different chromosomes using
the above criteria, we then attempted to design amplicons to type the SNPs at a subset of the sites
within each window.  We chose all sites with $|d|>0.975$, as well as the 8 SNPs on each chromosome
with the highest values of $|d|$, yielding 61 candidate SNPs (Figure~\ref{fig:wrap-candi}).

@@

%% Note to self.  The next section is documented in the project at:
%% /Users/eriq/Documents/work/assist/anthony_clemento/WRAP-crunching
These 48 amplicons were amplified and sequenced in 192 fish---96 Feather River spring run
and 96 winter run---on a MiSeq sequencer, and the variants were called using
GATK.  The resulting VCF file was used to estimate allele frequency differences between
winter run and Feather River spring run.  We also processed the sequence data using
the R package,
'microhaplot' (\url{https://github.com/ngthomas/microhaplot}), and visually inspected
loci for consistent allele depth ratio and numbers of haplotypes.  We then chose 24 amplicons for further
testing on the basis of allele frequency differences between Feather River spring run and
winter run, number of haplotypes, and ease of scoring.  These 24 markers were typed on
a variety of fish over the course of a year, and we finally chose three to include in
our California Chinook reference baseline: one amplicon on each of chromosomes
8, 12, and 16.  The estimated frequencies of all the alleles present in the reference baseline
in those three amplicons shows that there are not fixed differences at these markers between
winter run and all other reporting units (Table~\ref{tab:wrap-freqs}).
